---
title: "05 DATA/STAT 234 Review"
author: "Ben Sunshine"
format: 
  html:
    embed-resources: true
---

## 1/28/24

```{r}
library(tidyverse)
library(here)
library(ggrepel)
theme_set(theme_minimal())
hpi_df <- read_csv(here("data/hpi-tidy.csv"))

```

## 5.1

```{r}
hpi_us <- hpi_df |> filter(Country == "United States of America")

## install.packages("ggrepel")
# ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
#   geom_point() +
#   geom_text_repel(data = hpi_us, aes(label = Country)) +
#   geom_point(data = hpi_us, size = 3, shape = 1) +
#   labs(title = "Countries with a Higher Ecological Footprint Tend to Have Citizens with Higher Wellbeing", ## add title
#     subtitle = "Wellbeing is on a 1-10 scale", ## add subtitle (smaller text size than the title)
#     caption = "Data Source: http://happyplanetindex.org/countries", ## add caption to the bottom of the figure
#     x = "Ecological Footprint", ## change x axis label
#     y = "Wellbeing") ## change y axis label
```

## 5.1 Exercises

Exercise 1:

1.  mutate(): Allows you to modify variables or create new ones.
2.  arrange(): Allows you to arrange a data set by a based off a column or many.
3.  select(): Allows you to select certain columns from a data set.
4.  slice(): Allows you to select specific rows from a data set.
5.  filter(): Given a specified condition, filter() allows you to filter rows in a data set that meet a condition.
6.  group_by(): Allows you to group the data by one or many variables for, typically for use in analyzing groups.
7.  summarise(); Given statistics that the user specifies for a each group in a data frame.

Exercise 2:

```{r}
# Review mutate(). Create a new variable in hpi_df that is equal to Wellbeing / Footprint.
hpi_df %>%
  mutate(wellbeing_to_footprint_ratio = Wellbeing/Footprint)
```

Exercise 3:

```{r}
#  Review mutate() and case_when(). Create a new variable in hpi_df that is equal to "80s" if LifeExp is in the 80’s, "70s" if LifeExp is in the 70s, and "Below 70" if LifeExp is less than 70.
hpi_df %>%
  mutate(LifeExpectancy_groups = case_when(
    (LifeExpectancy >= 80) & (LifeExpectancy < 90) ~ "80s",
    (LifeExpectancy >= 70) & (LifeExpectancy < 80) ~ "70s",
    (LifeExpectancy < 70) ~ "Below 70"
  ))
```

Exercise 4:

```{r}
# Exercise 4. Review arrange(). Sort the hpi_df data so that the country with the highest LifeExp is in the first row and the country with the lowest LifeExp is in the last row.
hpi_df %>%
  arrange(desc(LifeExpectancy))
```

Exercise 5:

```{r}
# Review select(). Make a data frame from hpi_df that only has the variables Country and LifeExp.
hpi_df %>%
  select(Country, LifeExpectancy)
```

Exercise 6:

```{r}
# Review slice(). Make a data frame from hpi_df that only has the first 5 rows.
hpi_df %>%
  slice(1:5)
```

# Exercise 7:

```{r}
# Review filter(). Make a data frame from hpi_df that only has countries in the "Western World" Region.
hpi_df %>%
  filter(Region == 'Western World')
```

# Exercise 8:

```{r}
# Review filter(). Make a data frame from hpi_df that only has countries with a LifeExp less than 55.
hpi_df %>%
  filter(LifeExpectancy < 55)
```

# Exercise 9:

```{r}
# Review group_by() and summarise(). Make a table of the number of countries in each Region. (Hint: recall that the n() function is the counting function in dplyr).
hpi_df %>%
  group_by(Region) %>%
  summarise(count = n())
```

# Exercise 10:

```{r}
# Review group_by() and summarise(). Make a table with the maximum LifeExp in each Region.
hpi_df %>%
  group_by(Region) %>%
  summarise(max_life_expectancy = max(LifeExpectancy))
```

