---
title: "08 STAT 213 Review"
author: "Ben Sunshine"
format: 
  html:
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

## 2/11/24

## 8.1

```{r}
library(tidyverse)
library(broom)
library(here)
theme_set(theme_minimal())
```

```{r}
coffee_df <- read_csv(here("data/coffee_ratings.csv"))
coffee_mod <- lm(total_cup_points ~ species + aroma + flavor +
                   sweetness + moisture,
   data = coffee_df)

summary(coffee_mod)
#> 
#> Call:
#> lm(formula = total_cup_points ~ species + aroma + flavor + sweetness + 
#>     moisture, data = coffee_df)
#> 
#> Residuals:
#>     Min      1Q  Median      3Q     Max 
#> -9.5132 -0.3705  0.0726  0.5610  5.5844 
#> 
#> Coefficients:
#>                Estimate Std. Error t value Pr(>|t|)    
#> (Intercept)     7.04039    0.77377   9.099  < 2e-16 ***
#> speciesRobusta  2.85365    0.26861  10.624  < 2e-16 ***
#> aroma           1.95188    0.14575  13.392  < 2e-16 ***
#> flavor          5.09440    0.14042  36.281  < 2e-16 ***
#> sweetness       2.23956    0.06553  34.173  < 2e-16 ***
#> moisture       -1.88033    0.67368  -2.791  0.00533 ** 
#> ---
#> Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
#> 
#> Residual standard error: 1.168 on 1333 degrees of freedom
#> Multiple R-squared:  0.8891, Adjusted R-squared:  0.8887 
#> F-statistic:  2137 on 5 and 1333 DF,  p-value: < 2.2e-16
```

```{r}
summary(coffee_mod)$coefficients["moisture", 4]
#> [1] 0.005327594
```

```{r}
tidy(coffee_mod)
#> # A tibble: 6 × 5
#>   term           estimate std.error statistic   p.value
#>   <chr>             <dbl>     <dbl>     <dbl>     <dbl>
#> 1 (Intercept)        7.04    0.774       9.10 3.23e- 19
#> 2 speciesRobusta     2.85    0.269      10.6  2.31e- 25
#> 3 aroma              1.95    0.146      13.4  1.82e- 38
#> 4 flavor             5.09    0.140      36.3  4.73e-201
#> 5 sweetness          2.24    0.0655     34.2  2.41e-184
#> 6 moisture          -1.88    0.674      -2.79 5.33e-  3


tidy(coffee_mod) |> select(p.value)
#> # A tibble: 6 × 1
#>     p.value
#>       <dbl>
#> 1 3.23e- 19
#> 2 2.31e- 25
#> 3 1.82e- 38
#> 4 4.73e-201
#> 5 2.41e-184
#> 6 5.33e-  3

tidy(coffee_mod) |> filter(term == "aroma")
#> # A tibble: 1 × 5
#>   term  estimate std.error statistic  p.value
#>   <chr>    <dbl>     <dbl>     <dbl>    <dbl>
#> 1 aroma     1.95     0.146      13.4 1.82e-38
```

```{r}
glance(coffee_mod)
#> # A tibble: 1 × 12
#>   r.squared adj.r.squared sigma statistic p.value    df logLik   AIC   BIC
#>       <dbl>         <dbl> <dbl>     <dbl>   <dbl> <dbl>  <dbl> <dbl> <dbl>
#> 1     0.889         0.889  1.17     2137.       0     5 -2105. 4224. 4260.
#> # ℹ 3 more variables: deviance <dbl>, df.residual <int>, nobs <int>
```

```{r}
augment(coffee_mod)
#> # A tibble: 1,339 × 12
#>   total_cup_points species aroma flavor sweetness moisture .fitted  .resid
#>              <dbl> <chr>   <dbl>  <dbl>     <dbl>    <dbl>   <dbl>   <dbl>
#> 1             90.6 Arabica  8.67   8.83        10     0.12    91.1 -0.537 
#> 2             89.9 Arabica  8.75   8.67        10     0.12    90.5 -0.538 
#> 3             89.8 Arabica  8.42   8.5         10     0       89.2  0.577 
#> 4             89   Arabica  8.17   8.58        10     0.11    88.9  0.114 
#> 5             88.8 Arabica  8.25   8.5         10     0.12    88.6  0.214 
#> 6             88.8 Arabica  8.58   8.42        10     0.11    88.9 -0.0411
#> # ℹ 1,333 more rows
#> # ℹ 4 more variables: .hat <dbl>, .sigma <dbl>, .cooksd <dbl>,
#> #   .std.resid <dbl>

augment_df <- augment(coffee_mod)
augment_df |> filter(.cooksd > 1)
#> # A tibble: 1 × 12
#>   total_cup_points species aroma flavor sweetness moisture .fitted .resid
#>              <dbl> <chr>   <dbl>  <dbl>     <dbl>    <dbl>   <dbl>  <dbl>
#> 1                0 Arabica     0      0         0     0.12    6.81  -6.81
#> # ℹ 4 more variables: .hat <dbl>, .sigma <dbl>, .cooksd <dbl>,
#> #   .std.resid <dbl>
```

```{r}
ggplot(data = coffee_df, aes(x = total_cup_points)) +
  geom_histogram(bins = 15, fill = "white", colour = "black")
```


```{r}
augment_df |> filter(.hat > 0.2)
#> # A tibble: 2 × 12
#>   total_cup_points species aroma flavor sweetness moisture .fitted .resid
#>              <dbl> <chr>   <dbl>  <dbl>     <dbl>    <dbl>   <dbl>  <dbl>
#> 1             59.8 Arabica   7.5   6.67      1.33     0.1    58.4    1.38
#> 2              0   Arabica   0     0         0        0.12    6.81  -6.81
#> # ℹ 4 more variables: .hat <dbl>, .sigma <dbl>, .cooksd <dbl>,
#> #   .std.resid <dbl>

augment_df |> filter(.std.resid > 3 | .std.resid < -3)
#> # A tibble: 25 × 12
#>   total_cup_points species aroma flavor sweetness moisture .fitted .resid
#>              <dbl> <chr>   <dbl>  <dbl>     <dbl>    <dbl>   <dbl>  <dbl>
#> 1             82.8 Arabica  8.08   8.17     10        0.12    86.6  -3.85
#> 2             82.4 Arabica  5.08   7.75     10        0.11    78.6   3.79
#> 3             82.3 Arabica  7.75   8.08      6.67     0.11    78.1   4.27
#> 4             80.7 Arabica  7.67   7.5       6.67     0       75.2   5.51
#> 5             80   Arabica  7.58   7.75     10        0       83.7  -3.71
#> 6             79.9 Arabica  7.83   7.67     10        0       83.8  -3.87
#> # ℹ 19 more rows
#> # ℹ 4 more variables: .hat <dbl>, .sigma <dbl>, .cooksd <dbl>,
#> #   .std.resid <dbl>

ggplot(data = augment_df |> filter(.fitted > 25),
       aes(x = .fitted, y = .resid)) +
  geom_point() 
```

```{r}
library(palmerpenguins)
penguins
#> # A tibble: 344 × 8
#>   species island   bill_length_mm bill_depth_mm flipper_length_mm body_mass_g
#>   <fct>   <fct>             <dbl>         <dbl>             <int>       <int>
#> 1 Adelie  Torgers…           39.1          18.7               181        3750
#> 2 Adelie  Torgers…           39.5          17.4               186        3800
#> 3 Adelie  Torgers…           40.3          18                 195        3250
#> 4 Adelie  Torgers…           NA            NA                  NA          NA
#> 5 Adelie  Torgers…           36.7          19.3               193        3450
#> 6 Adelie  Torgers…           39.3          20.6               190        3650
#> # ℹ 338 more rows
#> # ℹ 2 more variables: sex <fct>, year <int>
```


Exercise 1. Examine the penguins data set in the palmerpenguins package: Fit a linear regression model with body_mass_g as the response variable and species and bill_length_mm as the predictors. Note that penguins with missing values for any of these three variables will be dropped from the analysis.

```{r}
penguin_mod <- lm(body_mass_g ~ species + bill_length_mm, data = penguins)
```

Exercise 2. Create a table of summary output, including coefficient estimates, standard errors, test statistics, and p-values, using one of the broom functions.

```{r}
broom::tidy(penguin_mod)
```


Exercise 3. Use glance() to glance at some of the relevant model statistics.

```{r}
glance(penguin_mod)
```

Exercise 4. Using augment(), create a plot of the residuals vs. the fitted values and evaluate the constant variance assumption.

```{r}
augment(penguin_mod) %>%
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point() 
```

-   There is no fanning or funneling of the residuals, but there are two groups which become visible, which indicate non constant variance.


Exercise 5. Using augment(), check to see if there are any penguins that are influential. Use 0.75 as your cut-off value for Cook’s distance.

```{r}
augment(penguin_mod) %>%
  filter(.cooksd >= 0.75)
```

-   There are no observations at or above the 0.75 threshold.


